\documentclass[12pt,a4paper]{report}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{authblk}
\usepackage{algorithm, algpseudocode}
\usepackage{geometry}
\usepackage{booktabs} 
\usepackage{longtable} 
\usepackage{array}


\newcommand{\ie}{\textit{i.e.}}
%customized todo
\definecolor{lightgreen}{rgb}{0.8,1.0,0.8}
\definecolor{red}{rgb}{0.8,0,0}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{blue}{rgb}{0,0,1}
\newcommand{\mytodo}[1]{\todo[inline,color=lightgreen]{TODO:#1}}
\newcommand{\myremark}[1]{\todo[inline, color=lightgreen]{\textbf{Remark:} #1}}

\title{The VNN-LIB standard for benchmarks\\2025}

\author[1]{Stefano Demarchi}
\author[2]{Dario Guidotti}
\author[2]{Luca Pulina}
\author[1]{Armando Tacchella}

\author[3]{Ann Roy}
\author[3]{Allen Antony}
\author[3]{Matthew Daggitt}

\affil[1]{University of Genoa, Viale Causa 13, 16145 Genoa, Italy}
\affil[2]{University of Sassari, Via Roma 151, 07100 Sassari, Italy}
\affil[3]{University of Western Australia, 35 Stirling Hwy, Crawley WA 6009, Australia}
  
\begin{document}

\maketitle

\begin{abstract}
This document presents VNN-LIB, a standard that formalises the query language for neural network verifiers. The standard uses the 
Open Neural Network Exchange (ONNX) format for model description and builds upon the Satisfiability Modulo Theories Library (SMT-LIB) 
format for query specification. Key among the standard is a formally defined syntax and semantics, complimentary tooling, as well as a 
command-line interface for verifiers. The goal is to foster greater robustness and interoperability in the neural network verification 
landscape.
\end{abstract}


\section*{Introduction}

As neural networks become increasingly integral to safety-critical and high societal impact applications\cite{1, 2, 3}, 
the need for robust verification of their properties is paramount. While neural networks have shown exceptional performance, 
they are susceptible to issues like adversarial examples\cite{3} and discriminatory behaviour\cite{4}.

Since 2023, the VNN-LIB standard\cite{5} has served as a de facto specification language for queries in the neural network verification 
community, notably supporting the annual VNN-COMP\cite{7}. It was conceived to facilitate the standardisation of solver interfaces and 
enable the collection of benchmarks in a common format. This document introduces a new version of the VNN-LIB standard which addresses 
shortcomings in expressiveness, conciseness, and formal rigour\cite{5, 7}.

This document is structured as follows. Chapter~\ref{sec:model} revisits the guidelines for model specification using ONNX.\@
Chapter~\ref{sec:specification_language} details the VNN-LIB query specification language, with Section~\ref{sec:smtlib} describing the 
SMT-LIB first-order logic specification language which VNN-LIB is inspired from, Section~\ref{sec:syntax} describing the formal syntax of VNN-LIB and 
Section~\ref{sec:semantics} discussing its semantics. Chapter~\ref{sec:query_categories} outlines different characteristics of queries, namely linearity, 
reachability, and normality. Chapter~\ref{sec:solver_interface} introduces a standard command line interface for verifiers. Finally, Chapter~\ref{sec:examples} 
presents examples illustrating how to use VNN-LIB to specify constraints.

\chapter{Network Representation}\label{sec:model}
\input{onnx_modelzoo_report}

\chapter{Query Specification Language}\label{sec:specification_language}

\section{SMT-LIB Property Language}\label{sec:smtlib}
\input{smtlib_property}

\section{Syntax}\label{sec:syntax}
\input{syntax}

\section{Semantics}\label{sec:semantics}

% ---------------------------------------------------------------

\chapter{Query Categories}\label{sec:query_categories}

\section{Linearity}\label{sec:linearity}

\section{Reachability}\label{sec:reachability}

\section{Disjunctive Normal Form}\label{sec:dnf}

% ---------------------------------------------------------------

\chapter{Solver Interface}\label{sec:solver_interface}
\section{Functionalities}\label{sec:functionalities}
\input{verifier_interface}

\section{Satisfying Assignments}\label{sec:satisfying_assignments}

% ---------------------------------------------------------------

\chapter{Examples}\label{sec:examples}

\appendix
\chapter{VNN-LIB LBNF Grammar}\label{app:lbnf_grammar}
\myremark{The full LBNF grammar, as developed using BNFC, should be included here.}

\begin{thebibliography}{9} 

\bibitem{1} 
D. Tang, B. Qin, and T. Liu, ``Document modelling with gated recurrent neural network for sentiment classification,'' in \emph{Proceedings of the 2015 Conference on Empirical Methods in Natural Language Processing}, 2015, pp. 1422--1432.

\bibitem{2} 
M. Bojarski, et al., ``End to end learning for self-driving cars,'' \emph{arXiv preprint arXiv:1604.07316}, 2016.

\bibitem{3} 
C. Szegedy, et al., ``Intriguing properties of neural networks,'' \emph{arXiv preprint arXiv:1312.6199}, 2013.

\bibitem{4} 
P. Zhang et al., ``White-box fairness testing through adversarial sampling,'' in \emph{2020 IEEE/ACM 42nd International Conference on Software Engineering (ICSE)}, New York, NY, USA:\@ ACM, 2020, pp. 949--960.\'doi: \href{https://doi.org/10.1145/3377811.3380331}{10.1145/3377811.3380331}.

\bibitem{5} 
S. Demarchi, D. Guidotti, L. Pulina, and A. Tacchella, ``Supporting Standardization of Neural Networks Verification with VNN-LIB and CoCoNet,'' in \emph{Proc. 6th Int. Workshop on Formal Methods for ML-Enabled Autonomous Systems (FoMLAS 2023)}, 2023, pp. 47--58.

\bibitem{6} 
C. Brix, S. Bak, C. Liu, and T. T. Johnson, ``The Fourth International Verification of Neural Networks Competition (VNN-COMP 2023): Summary and Results,'' 2023, doi: \href{https://doi.org/10.48550/arxiv.2312.16760}{10.48550/arxiv.2312.16760}.

\bibitem{7} 
L. C. Cordeiro et al., ``Neural Network Verification is a Programming Language Challenge,'' 2025, doi: \href{https://doi.org/10.48550/arxiv.2501.05867}{10.48550/arxiv.2501.05867}.

\bibitem{8} 
M. Forsberg and A. Ranta, ``The Labelled BNF Grammar Formalism,'' Department of Computing Science, Chalmers University of Technology and the University of Gothenburg, Gothenburg, Sweden, Feb. 11, 2005. [Online]. Available: \url{https://bnfc.digitalgrammars.com/LBNF-report.pdf}

\end{thebibliography}

\end{document}
